---
title: "My Sophisticated Analysis"
author: "Lo"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(ggplot2)
library(plotly)
```

## Here will be my really eye-catching title {data-width="650"}

### Temporary Introduction Tab Title

First Track

This is my own music. I composed it early 2024 for a saxophone quartet (soprano, alto, tenor, bass). It is the first movement of a larger 10 minute piece called The Water Bellows. It follows the story of a ship which meets an unfortunate end, inspired by the real-life HMAS Sydney II tragedy whose memorial I visited in Geraldton, Australia. This recording is an extract from the performance played in the Eindhoven concert hall.

Second Track

Also my own music. Composed in early 2024 as well for a theater collective in Tilburg called EELT. They did a collaboration with the women’s emancipation center in Tilburg called Feniks and I was lucky enough to be asked to write an atmospheric piece to accompany their theater and dance performance. The sounds used in the piece are recorded audio from instruments but also locations, such as train stations. They were then processed in a DAW.

### Here is my class corpus visualisation

```{r}
compmus2025 <- read_csv("C:/Users/Gebruiker/Downloads/compmus2025.csv")
compmus2025 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = arousal,
      size = danceability,
      colour = tempo
    )
  ) +
  geom_point() +  # Scatter plot.
  geom_smooth() +
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
    scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 6.5),
    breaks = c(3, 4, 5, 6, 7), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 8),
    breaks = c(3, 4, 5, 6, 7, 8),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Valence",
    y = "Arousal",
    colour = "Tempo"
  )
ggplotly()
```

------------------------------------------------------------------------

Here I will add a text to locate my tracks and analyze this upwards trend.

### Temporary Conclusion Tab Title

Here will be my final conclusion on my visualisations.
