---
title: "The Digital Against the Acoustic"
output: 
  flexdashboard::flex_dashboard:
    theme: 
      version: 4
      bootswatch: minty
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(ggplot2)
library(plotly)
library(rjson)
source("compmus.R")

major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.0, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
minor_key <-
  c(5.0, 2.0, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

```

Introduction {.storyboard}
=========================================

### Introducing my own tracks

First Track

This is my own music. I composed it early 2024 for a saxophone quartet (soprano, alto, tenor, bass). It is the first movement of a larger 10 minute piece called The Water Bellows. It follows the story of a ship which meets an unfortunate end, inspired by the real-life HMAS Sydney II tragedy whose memorial I visited in Geraldton, Australia. This recording is an extract from the performance played in the Eindhoven concert hall.

Second Track

Also my own music. Composed in early 2024 as well for a theater collective in Tilburg called EELT. They did a collaboration with the women’s emancipation center in Tilburg called Feniks and I was lucky enough to be asked to write an atmospheric piece to accompany their theater and dance performance. The sounds used in the piece are recorded audio from instruments but also locations, such as train stations. They were then processed in a DAW.

### Here is my class corpus visualisation

```{r}
compmus2025 <- read_csv("C:/Users/Gebruiker/Downloads/compmus2025.csv")
compmus2025 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = arousal,
      size = tempo,
      colour = danceability
    )
  ) +
  geom_point() +  # Scatter plot.
  geom_smooth() +
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
    scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 6.5),
    breaks = c(3, 4, 5, 6, 7), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 8),
    breaks = c(3, 4, 5, 6, 7, 8),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Valence",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```

------------------------------------------------------------------------

Here I will add a text to locate my tracks and analyze this upwards trend.

Chromagram {.storyboard}
=========================================

### Here is a chromagram for my first track

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

What is most interesting to note here is that my piece starts with an ostinato of G, clearly being shown in the chromagram. We can also see that the tonality remains in the same key where the I chord of the key (Gmajor) is dominant (G, B, D). We also see mediantics at play with the E minor chord being present.G# is present sometimes in the second half of the piece. I am wondering whether it could be that as the instruments warmed up when they played the piece, that the tuning went up and therefore G could get mistaken for G#.

### Here is a chromagram for my second track

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This is super interesting because this track is mostly atmospheric sounds recorded from a train station. This is reflected in the chromagram wonderfully. The only notes that are shown are when the instrument comes in in the middle. This chromagram almost shows structure better than pitch as three sections are clearly denoted (before the instrument, the instrument, after the instrument). when I played the instrument, I mostly taped a bunch of different notes but it seemed that this was interpreted here as primarily a G#. I wonder if this is because most of the notes had overtones that evoked G# too.

SSM {.storyboard}
=========================================

### Here is a Self-Similarity Matrix for Chroma - Track 1

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

I find this interesting because the piece does include a lot of repeats yet the graph tells us that it composes of a lot of new notes.The beginning and end are clearly dark. This is because the beginning has that G ostinato and the end of the piece includes repeating open fifths thrice.

### Here is a Self-Similarity Matrix for Timbre - Track 1

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This makes a lot of sense. The piece is a sax quartet so there is not much that changes in terms of timbre. Here we can again also see that ostinato on G at the beginning, highlighted by the completely similarity of that one note repeating as that section is black at the start. There is some lighter sections, possibly when the other saxophones start joining. In the second half, they all play all the time which is why it then settles into dark quickly.

### Here is a Self-Similarity Matrix for Chroma - Track 2

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

Here we see the same thing happen as with the chromagram where the novelty in yellow comes in when the instrument starts. At around 150, there is a darker chunk. I have not been able to find in the audio what causes this similarity to happen. I wonder if I maybe directly copied a part of my piece to this section and that I've forgotten that I've done it but the computer recognises this segment from earlier.

### Here is a Self-Similarity Matrix for Timbre - Track 2

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This matrix is intriguing because it doesn't see any of the various sampled outside train sounds as different from one another. It seems because they're all mechanical sounding or in the same range of pitch that they resemble one another. Funnily enough, in the same place as the dark square from the SSM for chroma, there's a lighter patch in the timbre SSM.

Chordogram and Keygram {.storyboard}
=========================================

### Chordogram for my first track

```{r}

"features/lo-l-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***
This is a chordogram of my first track. This second track is built with sampled outside sounds so it doesn't make much sense to think of it in terms of chords and keys. I believe we have seen enough information with the chroma and SSM of the second track to understand that nothing fruitful would come.

Here we see that a lot of chords that include the note G are selected, whilst those who don't have the note G as a major note in the chord are yellow. This must be because of the G ostinato at the beginning that the soprano saxophone holds. The beginning has three chord classifications: G, c, E (whether 7th chords, a minor chord, or a major chord). Eb chords are also prevalent, which explain why Cm is higlighted. I wrote this piece without a key, mostly just building chords around the note G so it makes sense to me for all the chords with a G to be present as I do rotate through a lot of weird clusters. 

The arpeggios later in the piece don't register as chords since the chordogram lacks the ability to group chroma by bar and have an overview of the general chord when the notes don't fall at the same time. The final long fifths at the end of my piece are the darkest part of the chordogram. Two chords are selected: E major and E minor. This makes sense as the chord is an open fifth so it could go either way and the chrodogram would need to check the previous context of the piece in order to decide whether a major or minor third would make more sense with the fifths.


### Keygram for my first track

```{r}
"features/lo-l-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***

The keygram seems to be very confused as to what the key could be. This makes sense, as I mentioned before, I wrote this without a key with mind. All I had decided was the theme and that the center of tonality was G. With the context that the piece isn't meant to have a key, I wonder if there is much information to be gleaned from this chart. I've also tried different norms and distances but it doesn't get more defined than this. This reflects well the nature of how I composed but does not give us much information.

Tempograms {.storyboard}
=========================================

### Here is a tempogram for my first track

```{r}
"features/lo-l-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

```

***

A tempogram for both my pieces doesn't really make much sense. This first piece, I originally wrote 60 bpm on my sheet music. Since there are a very wide range of time signature within this first movement, this leads to a poor result. The time signatures in my piece are 4/4, 1/8, 3/8, 5/8, 2/4, and 7/8. I think maybe it could pick up on the regular 8th notes that time all these time signature together, but there are many tuplets and polyrhythms blurring this 8th note underlying pulse.

The graph shows a widly inconcsistent tempo. This could have some sense since this is performed by real people and they might fluctuate. Something I should mention is that I did give my players the liberty to take time in between phrases for breathing since this piece requires a lot of continous playing for all of them. This can also mess with the tempo perception.

Overall, you can see these regular tempo harmonics reflected in the piece. The first line seems to sit at 100 bpm, quite higher than the bmp on my sheet music.

### Here is a tempogram for my second track

```{r}
"features/lo-l-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = FALSE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

```

***

These piece was designed without a time signature. Since it is a collection of outside recorded sounds, they do naturally have a rhythmic percussive sound which might create these onsets. However, there is nothing connecting all these sounds to one another. Every sound comes in randomly, the template I used in my DAW worked with seconds and not bars. I find it interesting that it found a tempo so high since the middle part with the instrument is quite slow. Of course the huge amount of reverb and delay I put on the instrument probably didn't help it understand whether it was a new onset or just the decay of the note.

This piece shows a more regular tempo overall compared to piece 1. It is electronic which could explain this a bit. I also use similar sounds and effects on the sounds throughout. It did find the timbre and chroma to not change at all, so it makes sense that it also interprets the tempo as staying the same. The tempogram seems to interpret the electronic sounds in a similar way as the chromagrams and SSMs. Here, again, you can see the doubling of the tempo octaves. It seems the lowest one sits at just around 70.


### Here is a cyclic tempogram for my first track

```{r}
"features/lo-l-1.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

```

***

The cyclic tempogram shows tempos that vary wildly from one second to the next rangin from 85bpm to 145bpm. At the very end, despite the three long chords I have composed, the tempogram interprets this as even higher that 145bpm, which is fascinating. It must be very confused by the sudden slowness that it compensates by subdividing. At 140 seconds, the tempo drops from very high to very low. Since the bulk of the middle part are arppegios following the 8th notes of the piece, I believe it is the time signature changes that gets it to be confused as to where the one is. This is why the tempo doesn't stay regular despite it being a very stable part of the piece. I'm very surprised with the cyclic tempogram that there is so much noise surrounding it since there isn't that much noise in the standard tempogram.


### Here is a tempogram for my first track

```{r}
"features/lo-l-2.json" |>
  compmus_tempogram(window_size = 8, hop_size = 1, cyclic = TRUE) |>
  ggplot(aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

```

***

This cyclic tempogram has a lot more variation that the standard tempogram. Around the 160 second time mark, there's a huge jump in tempo to 144. It looks like it lines up with the timbre and chroma difference at that time mark. I still have not been able to piece together what could be catching the attention of the program. You can also see the structure in this tempogram, the three parts this piece seems to be divided into where it is before the instrument, during, and after.


Conclusion
=========================================

### What conclusion can we draw from this?

Here will be my final conclusion on all my visualisations.
