---
title: "The Digital Against the Acoustic"
author: "Lo"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
---

```{r setup, include=FALSE}
library(tidyverse)
library(flexdashboard)
library(ggplot2)
library(plotly)
library(rjson)
source("compmus.R")

major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(5.0, 2.0, 3.5, 2.0, 4.5, 4.0, 2.0, 4.5, 2.0, 3.5, 1.5, 4.0)
minor_key <-
  c(5.0, 2.0, 3.5, 4.5, 2.0, 4.0, 2.0, 4.5, 3.5, 2.0, 1.5, 4.0)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

```

## The Digital Against the Acoustic {data-width="650"}

### Temporary Introduction Tab Title

First Track

This is my own music. I composed it early 2024 for a saxophone quartet (soprano, alto, tenor, bass). It is the first movement of a larger 10 minute piece called The Water Bellows. It follows the story of a ship which meets an unfortunate end, inspired by the real-life HMAS Sydney II tragedy whose memorial I visited in Geraldton, Australia. This recording is an extract from the performance played in the Eindhoven concert hall.

Second Track

Also my own music. Composed in early 2024 as well for a theater collective in Tilburg called EELT. They did a collaboration with the women’s emancipation center in Tilburg called Feniks and I was lucky enough to be asked to write an atmospheric piece to accompany their theater and dance performance. The sounds used in the piece are recorded audio from instruments but also locations, such as train stations. They were then processed in a DAW.

### Here is my class corpus visualisation

```{r}
compmus2025 <- read_csv("C:/Users/Gebruiker/Downloads/compmus2025.csv")
compmus2025 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = arousal,
      size = tempo,
      colour = danceability
    )
  ) +
  geom_point() +  # Scatter plot.
  geom_smooth() +
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
    scale_x_continuous(         # Fine-tune the x axis.
    limits = c(3, 6.5),
    breaks = c(3, 4, 5, 6, 7), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(3, 8),
    breaks = c(3, 4, 5, 6, 7, 8),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Valence",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```

------------------------------------------------------------------------

Here I will add a text to locate my tracks and analyze this upwards trend.

### Here is a chromagram for my first track

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

What is most interesting to note here is that my piece starts with an ostinato of G, clearly being shown in the chromagram. We can also see that the tonality remains in the same key where the I chord of the key (Gmajor) is dominant (G, B, D). We also see mediantics at play with the E minor chord being present.G# is present sometimes in the second half of the piece. I am wondering whether it could be that as the instruments warmed up when they played the piece, that the tuning went up and therefore G could get mistaken for G#.

### Here is a chromagram for my second track

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_chroma(norm = "manhattan") |>                 # Change the norm
  ggplot(aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This is super interesting because this track is mostly atmospheric sounds recorded from a train station. This is reflected in the chromagram wonderfully. The only notes that are shown are when the instrument comes in in the middle. This chromagram almost shows structure better than pitch as three sections are clearly denoted (before the instrument, the instrument, after the instrument). when I played the instrument, I mostly taped a bunch of different notes but it seemed that this was interpreted here as primarily a G#. I wonder if this is because most of the notes had overtones that evoked G# too.

### Here is a SSM Chroma 1

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

I find this interesting because the piece does include a lot of repeats yet the graph tells us that it composes of a lot of new notes.The beginning and end are clearly dark. This is because the beginning has that G ostinato and the end of the piece includes repeating open fifths thrice.

### Here is a SSM Timbre 1

```{r}
"features/lo-l-1.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This makes a lot of sense. The piece is a sax quartet so there is not much that changes in terms of timbre. Here we can again also see that ostinato on G at the beginning, highlighted by the completely similarity of that one note repeating as that section is black at the start. There is some lighter sections, possibly when the other saxophones start joining. In the second half, they all play all the time which is why it then settles into dark quickly.

### Here is a SSM Chroma 2

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_chroma(norm = "euclidean") |>                 # Change the norm
  compmus_self_similarity(
    feature = pc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

Here we see the same thing happen as with the chromagram where the novelty in yellow comes in when the instrument starts. At around 150, there is a darker chunk. I have not been able to find in the audio what causes this similarity to happen. I wonder if I maybe directly copied a part of my piece to this section and that I've forgotten that I've done it but the computer recognises this segment from earlier.

### Here is a SSM Timbre 2

```{r}
"features/lo-l-2.json" |>                           # Change the track
  compmus_mfccs(norm = "euclidean") |>                  # Change the norm
  compmus_self_similarity(
    feature = mfcc,
    distance = "cosine"                             # Change the distance
  ) |>   
  ggplot(aes(x = xtime, y = ytime, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()                                      # Change the theme?
```

------------------------------------------------------------------------

This matrix is intriguing because it doesn't see any of the various sampled outside train sounds as different from one another. It seems because they're all mechanical sounding or in the same range of pitch that they resemble one another. Funnily enough, in the same place as the dark square from the SSM for chroma, there's a lighter patch in the timbre SSM.

### Chordogram

```{r}

"features/lo-l-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    chord_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***
This is a chordogram of my first track. This second track is built with sampled outside sounds so it doesn't make much sense to think of it in terms of chords and keys. I believe we have seen enough information with the chroma and SSM of the second track to understand that nothing fruitful would come.

Here we see that a lot of chords that include the note G are selected, whilst those who don't have the note G as a major note in the chord are yellow. This must be because of the G ostinato at the beginning that the soprano saxophone holds. The beginning has three chord classifications: G, c, E (whether 7th chords, a minor chord, or a major chord). Eb chords are also prevalent, which explain why Cm is higlighted. I wrote this piece without a key, mostly just building chords around the note G so it makes sense to me for all the chords with a G to be present as I do rotate through a lot of weird clusters. 

The arpeggios later in the piece don't register as chords since the chordogram lacks the ability to group chroma by bar and have an overview of the general chord when the notes don't fall at the same time. The final long fifths at the end of my piece are the darkest part of the chordogram. Two chords are selected: E major and E minor. This makes sense as the chord is an open fifth so it could go either way and the chrodogram would need to check the previous context of the piece in order to decide whether a major or minor third would make more sense with the fifths.


### Keygram

```{r}
"features/lo-l-1.json" |> 
  compmus_chroma(norm = "identity") |> 
  compmus_match_pitch_templates(
    key_templates,         # Change to chord_templates if desired
    norm = "euclidean",       # Try different norms (and match it with what you used in `compmus_chroma`)
    distance = "cosine"   # Try different distance metrics
  ) |>
  ggplot(aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               # Change the colours?
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
```

***

The keygram seems to be very confused as to what the key could be. This makes sense, as I mentioned before, I wrote this without a key with mind. All I had decided was the theme and that the center of tonality was G. With the context that the piece isn't meant to have a key, I wonder if there is much information to be gleaned from this chart. I've also tried different norms and distances but it doesn't get more defined than this. This reflects well the nature of how I composed but does not give us much information.


### Temporary Conclusion Tab Title

Here will be my final conclusion on my visualisations.
